---
output: html_fragment
---

```{r, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
devtools::load_all()
options(width=600)
```

### Uploading Custom Networks

*Bieulergy* optimizes performance by representing networks as node/edge-lists and pre-computing various graph statistics used throughout the application. User data can be uploaded in the form of a serialized R object file `.rds`. The object must be a *named list of one or more interactive network objects*. 

Please refer to our [documentation](https://montilab.github.io/bieulergy/articles/docs/data.html) for creating `interactive network objects`.

*Bieulergy* assumes the following:  

- Networks are represented as *gene x gene* adjacency matrices with edges present in the upper triangle
- Networks are **undirected**
- Networks are the same dimensions and share the same row/column names

### Example

Here is an example of how the data should be formatted with 5 simulated networks of 100 genes.

```{r}
upload <- readRDS(file.path(system.file("extdata", package="bieulergy"), "yeast-networks.rds"))
```

A list of interactive omics objects...
```{r}
is(upload)
```

```{r}
sapply(upload, is)
```

Some pre-computed graph measures...
```{r}
print(upload$`Yeast #1`$properties)
head(upload$`Yeast #1`$nodes)
```
