---
title: "Interactive Network Objects"
vignette: >
    %\VignetteIndexEntry{Interactive Network Objects}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(comment="", cache=FALSE, fig.align="center", message=FALSE, warning=FALSE)
devtools::load_all(".")
```

Working with interactive network objects makes it easy to perform interactive analyses in Shiny and continue in the command line for extending analysis where you left off.

```{r}
yeast.networks <- readRDS(file.path(system.file("extdata", package="bieulergy"), "yeast-networks.rds"))
```

# Using Interactive Omics Network Objects 

Network objects are essentially undirected igraph objects with pre-computed graph and node level properties. Additionally, they allow a mapping of unique node identifiers to human-readable symbols. These symbols are typically gene or protein symbols. This allows for integrating enrichment-based analysis or incorporating node-level information into network analyses.

```{r}
network <- yeast.networks[[1]]

print(network)
```

## Object Structure

```{r}
str(network)
```

```{r}
head(network$nodes)
head(network$edges)
```

### Converting unique nodes identifiers to mapped symbols

```{r}
network$get.symbols(c("749", "109", "740"))
```

### Find interactors or filter sub networks

```{r}
network$get.interactors(c("109"))
network$get.interactors(c("109", "107"))
```

```{r}
network$plt.subnetwork(ids=c("109", "107"))
network$plt.subnetwork(ids=c("109", "107"), indirect.edges=TRUE)
```

```{r}
network$plt.subnetwork(ids=c("109", "107"), degree=3)
```

```{r}
network$plt.subnetwork(ids=c("109", "107"), degree=3, use.symbols=TRUE)
```

```{r}
network$plt.subnetwork(ids=c("109", "107"), degree=3, use.symbols=TRUE, node.color="degree")
```

```{r}
network$plt.subnetwork(ids=c("109", "107"), degree=5, use.symbols=TRUE, node.color="degree", layout="layout_on_grid")
```

```{r}
network$plt.subnetwork(ids=c("109", "107", "607"), degree=0)
```

### Omics Accessory Data

What if we have lots of multi-omics data we'd like to incorporate into the node/edge properties?

```{r}
network$plt.subnetwork(ids=c("285"),
                       degree=2,
                       use.symbols=TRUE,
                       node.size="eigen",
                       node.color=list("lfc_mrna", cfx(c("blue", "white", "red"))),
                       node.border=list("snp_frq", cfx(c("#e9e9e9", "black"))),
                       node.shape="is_tf")
```

### Sub Network Enrichment

What if we wanted to test enrichment of biological pathways within a sub network?

```{r}
library(hypeR)
genesets <- enrichr_download(genesets="KEGG_2019", db="YeastEnrichr")

signature <- network$get.interactors(ids=c("285"), degree=2, remove.ids=FALSE, use.symbols=TRUE)
head(signature)

hyp <- hypeR(signature, genesets, background=16454)
hyp_dots(hyp)

head(hyp$data[,2:6])
```
